%\chapter{Flow Topo}
\section{Flow Topo}
\label{sec:flow-topo}


\subsection{Description}
%-----------------------------------------------------------------------------
Extracts critical points from velocity data, in the sense of vector field topology.
 

\subsection{Input}
%-----------------------------------------------------------------------------
\begin{itemize}
\item
  data grid (unstructured)
  \begin{itemize}
  \item
    velocity (3-vect)
  \item
    wall distance (optional)
  \end{itemize}
\end{itemize}


\subsection{Output}
%-----------------------------------------------------------------------------
\begin{itemize}
\item
  point type geometry / 1D array of 3D points (the critical points)
\end{itemize}


\subsection{Parameters}
%-----------------------------------------------------------------------------
\begin{itemize}

\item
  \textbf{divide by walldist}: if active, velocity is divided by wall distance. This results in a vector field that is proportional to the wall shear stress at the boundaries.

\item
  \textbf{interior critical points}: if active, critical points are extracted in the interior of the flow.

\item
  \textbf{boundary critical points}: if active, critical points are extracted at the boundary of the flow.

\item
  \textbf{generate seeds}: if active, seeds are generated at each critical point. The seeds are generated on two circles of radius \emph{radius}, \emph{seeds per circle} seeds are generated on each circle. The two circles are parallel to the 2D manifold of the critical point. The centers of the circles are offset by \emph{offset} from the critical point, along the direction of the 1D manifold.

\item
  \textbf{seeds per circle}: number of seeds per circle.

\item
  \textbf{radius}: radius of the circles.

\item
  \textbf{offset}: offset of the circle centers from the critical point.

\end{itemize}


\subsection{Implementation}
%-----------------------------------------------------------------------------


\subsubsection{Version}
%.............................................................................

2007-01-11


\subsubsection{Author}
%.............................................................................

Filip Sadlo, Ronald Peikert


\subsection{See Also}
%-----------------------------------------------------------------------------


\subsubsection{Related Modules}
%.............................................................................


